---
title: "Take-home_Ex1"
date: "January 13,2024"
date-modified: "last-modified"
execute: 
  eval: true 
  echo: true
  warning: false
editor: visual 
---

# Project Brief

"Singapore managed to achieve excellence without wide differences between children from wealthy and disadvantaged families," said OECD education director Andreas Schleicher in a BBC piece. However, the general public, however, is adamant that there are still differences between educational attainment and socioeconomic position.

This project will explore multiple possible factors that could help to verify and investigate the statement further, utilizing the proper R statistical graphics techniques, both interactive and static.

# Project Objective

This project aims to assess if there still social difference among different background of families. Students' performance on math, reading and science will be the main metrics. The effects brought by school, gender and socioeconomic status on student's academic performance will be tested in this project. Besides, the socioeconomic status will be demonstrated through number of digital devices, number of musical instruments, and food availability.

# 1. Data Preparation

## 1.1 Loading R package

Let's load the libraries and R package needed for this assignment using `pacman::p_load()` function:

```{r}
pacman::p_load(tidyverse, haven, patchwork)
```

The loaded packages are:

-   tidverse

-   haven:

-   ggrepel:

-   patchwork:

-   ggthemes:

-   hrbrthemes:

-   ggrain:

## 1.2 Dataset

The dataset using in this assignment is taking from [PISA 2022 database](https://www.oecd.org/pisa/data/2022database/), called student questionnaire data file. It is uploaded as stu_qqq.

### 1.2.1 Import the datasets

In terms of import dataset, there is an extra step before getting the final dataset. Since we are only interest in Singapore students' related performance and information, a filter is applied to extract all Singapore data.

```{r}
#| eval: false
stu_qqq <-read_sas("data/cy08msp_stu_qqq.sas7bdat")
```

This is the output dataset using `filter ()` function to select all Singapore student data.

```{r}
#| eval: false
stu_qqq_SG <-stu_qqq %>%
  filter(CNT == "SGP")
```

```{r}
#| eval: false
write_rds(stu_qqq_SG,
          "data/stu_qqq_SG.rds")
```

```{r}
stu_qqq_SG <- 
read_rds("data/stu_qqq_SG.rds")
```

### 1.2.2 Summary Statistics of Stu_qqq_data

This is a summary statistics of the stu_qqq data, after using `glimpse ()` to get a overview of basic information of the dataset (i.e., number of records, detail of variables ), applying `duplicated ()` and `is.na()` to check if there is any duplicates data and missing values.

After running the function, here is the results:

::: panel-tabset
## Initial Raw Dataset Overview

```{r}
dim_desc(stu_qqq_SG)
head(stu_qqq_SG,5)
```

-   Observations: 6606

-   Columns: 1279

The dataset contains basic information of sudents (i.e., school ID, birth month and year), and their answer for questions under different questionnaire.

## Data Checking

Duplicates

```{r}
stu_qqq_SG[duplicated(stu_qqq_SG),]
```

-   There is no duplicates inside the dataset

Missing Values

```{r}
sum(is.na(stu_qqq_SG))
```

-   There are 4168500 values missing, missing value are represented NA in the dataset.

-   There are significant number of missing value in the dataset, however these are spread across the dataset, not all columns that are used in EDA will contain these amount of missing value.
:::

# 2. Data Wrangling

## 2.1 Select Useful Columns

Since there are quite number of variables, it is not realistic if we analyze all the variables. Therefore, the variable select below are the ones that this assignment will focus on, using `select ()` .

Besides, the original variable is named in a code base and numeric way, so it is easy for people to understand and use if it is named as word. The name of variable below is after using `remame()` .

```{r}
stu_qqq_eda<-stu_qqq_SG %>%
  select(CNTSCHID,CNTSTUID,ST004D01T,ST258Q01JA,PV1MATH,PV1READ, PV1SCIE,ESCS)%>%
  rename(c(Gender="ST004D01T", 
           School_ID="CNTSCHID", 
           Food_Availability='ST258Q01JA',
           Student_ID='CNTSTUID',
           Math_Score= "PV1MATH",
           Reading_Score="PV1READ",
           Science_Score="PV1SCIE"))
```

Stu_qqq_eda is the final dataset which will be used in this assignment with 6606 observations of 9 variables.

## 2.2 Change Data Types

School_ID, Gender, Num_DigitaDevice, Food_Availability, Num_MusicalInstruments are categorized as num, but they are categorical data with different levels. Therefore, they are cast as factor type using `as.factor()`.

```{r}
stu_qqq_eda$School_ID = as.factor(stu_qqq_eda$School_ID)
stu_qqq_eda$Gender = as.factor(stu_qqq_eda$Gender)
stu_qqq_eda$Food_Availability = as.factor(stu_qqq_eda$Food_Availability)
```

## 2.3 Recode variable

Currently, "gender" variable is using "1", "2" to represent for male and female. However, it might be not clear in the future analysis for people to quickly identify female and male. Therefore, rename the option is important for people to get a quick understanding, using `recode()` function.

```{r}
stu_qqq_eda <- stu_qqq_eda %>%
  mutate(Gender=(recode(Gender, '1'="Female", '2'="Male")),
         Food_Availability=(recode(Food_Availability,
                                   '1'='never',
                                   '2'='once a week',
                                   '3'='2-3 a week',
                                   '4'='4-5 a week',
                                   '5'='everyday')))

```

## 2.4 Final dataset overview

```{r}
glimpse(stu_qqq_eda)
```

# 3. Exploratory Data Analysis (EDA)

Before conduct EDA, since the target metrics is students performance in math, reading and science, it is necessary to see the overall distribution of students performance across all this three subjects.

## Students Performance Distribution

Histogram are used to show the frequency distribution for numerical data. `geom_histogram()` function in `ggplot` is applied to draw the graph.

::: panel-tabset
## Math

```{r}
#| code-fold: true
#| code-summary: "Show code"
ggplot(data=stu_qqq_eda, 
       aes(x=Math_Score)) +
  geom_histogram(color= "grey10",
           fill= '#FFC0CB') +
  labs(x= "Math_Score",
       y="No.of students",
       title= "Distribution of Student Math Performance") +
  geom_vline(aes(xintercept=median(stu_qqq_eda$Math_Score, na.rm = T)), 
             color="#595DE5", 
             linewidth=1, 
             linetype="dashed") + 
  annotate("text", 
           x=675, 
           y=70, 
           label= paste("Median =", round(median(stu_qqq_eda$Math_Score, na.rm = T), 3)), 
           color= "#595DE5", 
           size=3.5) +
  geom_vline(aes(xintercept=mean(stu_qqq_eda$Math_Score, na.rm = T)),
             colour="red",
             linewidth=1,
             linetype="dashed") +
  annotate("text",
           x=500,
           y=20,
           label=paste("Mean =", round(mean(stu_qqq_eda$Math_Score, na.rm = T), 3)),
           color="red",
           size=3.5) +
  theme(plot.background=element_rect(fill="#f5f5f5",colour="#f5f5f5"))
```

## Reading

```{r}
#| code-fold: true
#| code-summary: "Show code"
ggplot(data=stu_qqq_eda, 
       aes(x=Reading_Score)) +
  geom_histogram(color= "grey10",
           fill= '#FFC0CB') +
  labs(x= "Reading_Score",
       y="No.of students",
       title= "Distribution of Student Reading Performance") +
  geom_vline(aes(xintercept=median(stu_qqq_eda$Reading_Score, na.rm = T)), 
             color="#595DE5", 
             linewidth=1, 
             linetype="dashed") + 
  annotate("text", 
           x=650, 
           y=60, 
           label= paste("Median =", round(median(stu_qqq_eda$Reading_Score, na.rm = T), 3)), 
           color= "#595DE5", 
           size=3.5) +
  geom_vline(aes(xintercept=mean(stu_qqq_eda$Reading_Score, na.rm = T)),
             colour="red",
             linewidth=1,
             linetype="dashed") +
  annotate("text",
           x=460,
           y=50,
           label=paste("Mean =", round(mean(stu_qqq_eda$Reading_Score, na.rm = T), 3)),
           color="red",
           size=3.5) +
  theme(plot.background=element_rect(fill="#f5f5f5",colour="#f5f5f5"))
```

## Science

```{r}
#| code-fold: true
#| code-summary: "Show code"
ggplot(data=stu_qqq_eda, 
       aes(x=Science_Score)) +
  geom_histogram(color= "grey10",
           fill= '#FFC0CB') +
  labs(x= "Science_Score",
       y="No.of students",
       title= "Distribution of Student Science Performance") +
  geom_vline(aes(xintercept=median(stu_qqq_eda$Science_Score, na.rm = T)), 
             color="#595DE5", 
             linewidth=1, 
             linetype="dashed") + 
  annotate("text", 
           x=650, 
           y=60, 
           label= paste("Median =", round(median(stu_qqq_eda$Science_Score, na.rm = T), 3)), 
           color= "#595DE5", 
           size=3.5) +
  geom_vline(aes(xintercept=mean(stu_qqq_eda$Science_Score, na.rm = T)),
             colour="red",
             linewidth=1,
             linetype="dashed") +
  annotate("text",
           x=480,
           y=50,
           label=paste("Mean =", round(mean(stu_qqq_eda$Science_Score, na.rm = T), 3)),
           color="red",
           size=3.5) +
  theme(plot.background=element_rect(fill="#f5f5f5",colour="#f5f5f5"))
```
:::

From the histogram show above, students' math, reading, and science demonstrated a normal distribution with a similar mean and median score across this three subjects. More than half of the students' score is in the range from 500-600.

## 3.1 Are there difference in student performance between gender ?

### Distribution

The first EDA is to explore the difference in students' performance between male and female.

From the bar chart below shows that number of students between male and female have no significant difference. There are around 3500 students for male and female.

```{r}
ggplot(data = stu_qqq_eda,
       aes(x = Gender)) +
  geom_bar()
```

Moving on to the analysis part, the idea of choosing histogram is that it can clearly show performance distribution among every range of score. In order to compare between gender clearly, using `scale_fill_manual()` function to construct the histogram together with male and female.

::: panel-tabset
## Math

```{r}
#| code-fold: true
#| code-summary: "Show code"
ggplot(data=stu_qqq_eda, 
       aes(x=Math_Score,
           fill = Gender)) +
  geom_histogram(bins = 10,
                 position= "dodge",
                 color="grey10") +
  scale_fill_manual(values = c("#F98B88", "#F5FFFA")) +
  scale_y_continuous(breaks = seq(0, 3200, by = 500)) +
  scale_x_continuous(breaks = seq(0, 1000, by = 100)) +
  ylim(c(0,1000)) +
  stat_bin(bins = 10,
           geom = "text",
           aes(label = after_stat(count)),
           vjust = -.5, size=2.5,
           position = position_dodge())
```

Based on the histogram output, we can see that both male and female represent normal distribution and there is no obvious difference on math score across every bin.

## Reading

```{r}
#| code-fold: true
#| code-summary: "Show code"
ggplot(data=stu_qqq_eda, 
       aes(x=Reading_Score,
           fill = Gender)) +
  geom_histogram(bins = 10,
                 position= "dodge",
                 color="grey10") +
  scale_fill_manual(values = c("#F98B88", "#F5FFFA")) +
  scale_y_continuous(breaks = seq(0, 3200, by = 500)) +
  scale_x_continuous(breaks = seq(0, 1000, by = 100)) +
  ylim(c(0,1000)) +
  stat_bin(bins = 10,
           geom = "text",
           aes(label = after_stat(count)),
           vjust = -.5, size= 2.5,
           position = position_dodge())
```

Based on the histogram output, we can see that both male and female represent normal distribution and there is no obvious difference on reading score across every bin.

## Science

```{r}
#| code-fold: true
#| code-summary: "Show code"
ggplot(data=stu_qqq_eda, 
       aes(x= Science_Score,
           fill = Gender)) +
  geom_histogram(bins = 10,
                 position= "dodge",
                 color="grey10") +
  scale_fill_manual(values = c("#F98B88", "#F5FFFA")) +
  scale_y_continuous(breaks = seq(0, 3200, by = 500)) +
  scale_x_continuous(breaks = seq(0, 1000, by = 100)) +
  ylim(c(0,1000)) +
  stat_bin(bins = 10,
           geom = "text",
           aes(label = after_stat(count)),
           vjust = -.5, size= 2.5,
           position = position_dodge())
```

Based on the histogram output, we can see that both male and female represent normal distribution and Males' science score slightly higher than females'. Overall, there is no dramatic difference of science score between gender.
:::

In conclusion, after observing the histogram among the three subjects and gender. There seems no significant difference on student's academic performance between male and female.

## 3.2 Are there difference in student performance among various schools?

This EDA is to explore the difference in students' performance among various school. Before plotting, the average of students' score is calculated using `mean ()` function for a overall demonstration among schools.

```{r}
stu_avg_score <- stu_qqq_eda %>%
  group_by(School_ID) %>%
  mutate(math_avg = mean(Math_Score),
         science_avg = mean(Science_Score),
         reading_avg = mean(Reading_Score)) %>%
  select(School_ID, math_avg, science_avg, reading_avg) %>%
  unique() 
```

::: panel-tabset
## Math

```{r}
ggplot(data=stu_avg_score,
       aes(x= math_avg)) +
  geom_boxplot(width=0.1) +
  geom_point(aes(y= -0.1), position= position_jitter(width = 0.1, height = 0), color = "red", alpha = 0.5) +
  ylim(c(-.1,.1)) +
  scale_y_continuous(NULL, breaks = NULL)
  labs(x= "Math Average among schools", y= NULL)

```

From the graph above, it shows that the score spread across from 400 to 800, which means that the variance is large. Therefore, it indicated that the difference of schools make a contribution towords students' math score performance.

## Science

```{r}
ggplot(data=stu_avg_score,
       aes(x= science_avg)) +
  geom_boxplot(width=0.1) +
  geom_point(aes(y= -0.1), position= position_jitter(width = 0.1, height = 0), color = "red", alpha = 0.5) +
  ylim(c(-.1,.1)) +
  scale_y_continuous(NULL, breaks = NULL)
  labs(x= "Science Average among schools", y= NULL)
```

From the graph above, it shows that the score spread across from 350 to 750 which means that the variance is large. Therefore, it indicated that the difference of schools make a contribution towards students' science score performance.

## Reading

```{r}
ggplot(data=stu_avg_score,
       aes(x= reading_avg)) +
  geom_boxplot(width=0.1) +
  geom_point(aes(y= -0.1), position= position_jitter(width = 0.1, height = 0), color = "red", alpha = 0.5) +
  ylim(c(-.1,.1)) +
  scale_y_continuous(NULL, breaks = NULL)
  labs(x= "Reading Average among schools", y= NULL)
```

From the graph above, it shows that the score spread across from 300 to 700, which means that the variance is large. Therefore, it indicated that the difference of schools make a contribution to students' reading score performance.
:::

In conclusion, from the output of score performance and schools, it generally demonstrated that school have a possibility to be an indicator to students' academic performance.

## 3.3 Socioeconomic Status

In this section, the main focus is related to students' socioeconomic status. Before go into the specific variable, lets first have a look to the distribution of their socioeconomic status

### Distribution of SocialEconomic Status Score

```{r}
ggplot(data=stu_qqq_eda, 
       aes(x=ESCS)) +
  geom_histogram(color= "grey10",
           fill= '#FFC0CB') +
  labs(x= "Socioeconomic status",
       y="No.of students",
       title= "Distribution of Socioeconomic Performance") +
geom_vline(aes(xintercept=median(stu_qqq_eda$ESCS, na.rm = T)), 
             color="#595DE5", 
             linewidth=1, 
             linetype="dashed") + 
  annotate("text", 
           x=1.5, 
           y=100, 
           label= paste("Median =", round(median(stu_qqq_eda$ESCS, na.rm = T), 3)), 
           color= "#595DE5", 
           size=3.5) +
  geom_vline(aes(xintercept=mean(stu_qqq_eda$ESCS, na.rm = T)),
             colour="red",
             linewidth=1,
             linetype="dashed") +
  annotate("text",
           x=-0.5,
           y=100,
           label=paste("Mean =", round(mean(stu_qqq_eda$ESCS, na.rm = T), 3)),
           color="red",
           size=3.5) +
  theme(plot.background=element_rect(fill="#f5f5f5",colour="#f5f5f5"))
```

From the histogram above, it demonstrated that the range of socioeconomic status is from -4 to 3.5 and it is slightly skewed to the right side.

## 3.3.1 Student performance and socioeconomic status

In this EDA, it will show the relationship between student academic score and socioeconomic status score using scatter plot with `cor()` function.

::: panel-tabset
## Math

```{r}
cor_Math <- round(cor(stu_qqq_eda$Math_Score, stu_qqq_eda$ESCS, use = "complete.obs"), 2)

ggplot(data = stu_qqq_eda,
       aes(y = Math_Score, x = ESCS)) +
  geom_point(size = 1)+
  geom_smooth(method = lm) + 
  annotate("text", x=0, y=500, label=paste0("r = ", cor_Math), color = 'green')
```

From the graph showing above, it shows that there is moderate positive relationship between students' math score and their socioeconomic status score. It indicated that if the students have higher socioeconomic status score, they will more likely to have a higher math score.

## Reading

```{r}
cor_Reading <- round(cor(stu_qqq_eda$Reading_Score, stu_qqq_eda$ESCS, use = "complete.obs"), 2)

ggplot(data = stu_qqq_eda,
       aes(y = Reading_Score, x = ESCS)) +
  geom_point(size = 1)+
  geom_smooth(method = lm) + 
  annotate("text", x=0, y=500, label=paste0("r = ", cor_Reading), color = 'green')
```

From the graph showing above, it shows that there is moderate positive relationship between students' reading score and their socioeconomic status score. It indicated that if the students have higher socioeconomic status score, they will more likely to have a higher reading score.

## Science

```{r}
cor_Science <- round(cor(stu_qqq_eda$Science_Score, stu_qqq_eda$ESCS, use = "complete.obs"), 2)

ggplot(data = stu_qqq_eda,
       aes(y = Science_Score, x = ESCS)) +
  geom_point(size = 1)+
  geom_smooth(method = lm) + 
  annotate("text", x=0, y=500, label=paste0("r = ", cor_Science), color = 'green')
```

From the graph showing above, it shows that there is moderate positive relationship between students' science score and their socioeconomic status score. It indicated that if the students have higher socioeconomic status score, they will more likely to have a higher science score.
:::

In conclusion, from all three scatter plot, a consistent pattern can be observed from the three subject with moderate positive relationship. It demonstrated that higher socioeconomic status indicated the better academic performance.

## 3.3.3 Will food availability affect student performance ?

In this EDA, it will show the difference of students' academic performance on the degree of food availability.

Here is a distributions of food availability:

### Distributions

```{r}
ggplot(data = stu_qqq_eda,
       aes(x = Food_Availability)) +
  geom_bar()
```

From the bar graph above, we can see that most of the students don't have concern of food availability.

`drop_na()` function is applied to remove NA data which is not useful in the analysis

```{r}
cleaned_stu_qqq_eda <- stu_qqq_eda %>%
  select(Math_Score, Science_Score, Reading_Score, Food_Availability) %>%
  drop_na()

```

::: panel-tabset
## Math

```{r}
ggplot(data = cleaned_stu_qqq_eda, aes(x = factor(Food_Availability), y = Math_Score)) +
  geom_violin() +
  labs(x = "Degree of Food Concern", y = "Math Score") +
  ggtitle("Distribution of Math Scores for Each level of Food Availability") +
  stat_summary(fun = "mean",
               geom = "point",
               color = "blue")

```

From the violin graph above, the mean of math score among different degree of food concern frequency is reducing from never to everyday. It indicated that a student who have higher math score more likely to have less food availability concern.

## Reading

```{r}
ggplot(data = cleaned_stu_qqq_eda, aes(x = factor(Food_Availability), y = Reading_Score)) +
  geom_violin() +
  labs(x = "Degree of Food Concern", y = "Reading Score") +
  ggtitle("Distribution of Reading Scores for Each level of Food Availability") +
  stat_summary(fun = "mean",
               geom = "point",
               color = "blue")

```

From the violin graph above, the mean of reading score among different degree of food concern frequency is reducing from never to everyday. It indicated that a student who have higher reading score more likely to have less food availability concern.

## Science

```{r}
ggplot(data = cleaned_stu_qqq_eda, aes(x = factor(Food_Availability), y = Science_Score)) +
  geom_violin() +
  labs(x = "Degree of Food Concern", y = "Science Score") +
  ggtitle("Distribution of Science Scores for Each level of Food Availability") +
  stat_summary(fun = "mean",
               geom = "point",
               color = "blue")
```

From the violin graph above, the mean of science score among different degree of food concern frequency is reducing from never to everyday. It indicated that a student who have higher science score more likely to have less food availability concern.
:::
