---
title: "Take-home_Ex1"
date: "January 13,2024"
date-modified: "last-modified"
execute: 
  eval: true 
  echo: true
  warning: false
editor: visual 
---

# Project Brief

"Singapore managed to achieve excellence without wide differences between children from wealthy and disadvantaged families," said OECD education director Andreas Schleicher in a BBC piece. However, the general public, however, is adamant that there are still differences between educational attainment and socioeconomic position, particularly between neighborhood and elite schools, as well as between immigrant and non-immigrant families and students from higher socioeconomic families.

This project aims to reveal the certain characteristics of student to explore the relation with their grade, utilizing the proper R statistical graphics techniques, both interactive and static.

# Project Objective

# 1: Data Preparation

## 1.1 Loading R package

The code chunck used is as follows:

```{r}
pacman::p_load(tidyverse,haven,ggrepel, patchwork, ggthemes, hrbrthemes)
```

## 1.2 Dataset

### 1.2.1 Import the datasets

```{r}
#| eval: false
stu_qqq <-read_sas("data/cy08msp_stu_qqq.sas7bdat")
```

```{r}
#| eval: false
stu_qqq_SG <-stu_qqq %>%
  filter(CNT == "SGP")
```

```{r}
#| eval: false
write_rds(stu_qqq_SG,
          "data/stu_qqq_SG.rds")
```

```{r}
stu_qqq_SG <- 
read_rds("data/stu_qqq_SG.rds")
```

### 1.2.2 Summary Statistics of Stu_qqq_data

::: panel-tabset
## Dataset Overview

```{r}
str(stu_qqq_SG)
```

```{r}
glimpse(stu_qqq_SG)
```

## Data Checking

Duplicates

```{r}
stu_qqq_SG[duplicated(stu_qqq_SG),]
```

Missing Values

```{r}
sum(is.na(stu_qqq_SG))
```
:::

### 1.2.3 A Quick Visual Look of Stu_qqq_data

```{r}

```

# 2. Data Wrangling

## 2.1 Select Useful Columns  

```{r}
stu_qqq_eda<-stu_qqq_SG %>%
  select(CNTSCHID,CNTSTUID,ST004D01T,ST001D01T,ST253Q01JA,ST258Q01JA, ST251Q06JA)%>%
  rename(c(Stu_Grade='ST001D01T', 
           Gender="ST004D01T", 
           School_ID="CNTSCHID", 
           Num_DigitalDevice='ST253Q01JA',
           Food_Availability='ST258Q01JA',
           Num_MusicalInstruments='ST251Q06JA', 
           Student_ID='CNTSTUID'))
```

## 2.2 Change Data Types 

```{r}
stu_qqq_eda$School_ID = as.factor(stu_qqq_eda$School_ID)
stu_qqq_eda$Gender = as.factor(stu_qqq_eda$Gender)
stu_qqq_eda$Num_DigitalDevice = as.factor(stu_qqq_eda$Num_DigitalDevice)
stu_qqq_eda$Food_Availability = as.factor(stu_qqq_eda$Food_Availability)
stu_qqq_eda$Num_MusicalInstruments= as.factor(stu_qqq_eda$Num_MusicalInstruments)
```

## 2.3 Recode variable 

```{r}
stu_qqq_eda <- stu_qqq_eda %>%
  mutate(Gender=(recode(Gender, '1'="Female", '2'="Male")))
```

# 3. Exploratory Data Analysis 

## 3.1 Students Performance 

## 3.1.1 Distribution 

```{r}
ggplot(data=stu_qqq_eda, 
       aes(x=Stu_Grade)) +
  geom_bar(color= "grey10",
           fill= '#FFC0CB') +
  labs(x= "Student Grade",
       y="No.of students",
       title= "Distribution of Student Performance") +
  geom_vline(aes(xintercept=median(stu_qqq_eda$Stu_Grade, na.rm = T)), 
             color="#595DE5", 
             linewidth=1, 
             linetype="dashed") + 
  annotate("text", 
           x=10.5, 
           y=6000, 
           label= paste("Median =", round(median(stu_qqq_eda$Stu_Grade, na.rm = T), 3)), 
           color= "#595DE5", 
           size=3.5) +
  geom_vline(aes(xintercept=mean(stu_qqq_eda$Stu_Grade, na.rm = T)),
             colour="red",
             linewidth=1,
             linetype="dashed") +
  annotate("text",
           x=9.5,
           y=5000,
           label=paste("Mean =", round(mean(stu_qqq_eda$Stu_Grade, na.rm = T), 3)),
           color="red",
           size=3.5) +
  theme(plot.background=element_rect(fill="#f5f5f5",colour="#f5f5f5"))
```

## 3.1.2 Are there difference in student performance between gender ? 

```{r}
ggplot(data=stu_qqq_eda, 
       aes(x=Stu_Grade,
           fill = Gender)) +
  geom_bar(position= "dodge",
           color="grey10") +
  scale_fill_manual(values = c("#F98B88", "#F5FFFA"))+
  scale_y_continuous(breaks = seq(0, 3200, by = 500)) +
  ylim(c(0,3500))+
  geom_text(stat="count", 
      aes(label=paste0(after_stat(count))),
      vjust=-0.5,
      position = position_dodge(width = 0.9))

```

## 3.1.3 Are there difference in student performance among various schools?

## 3.1.4 Are there difference in student performance between gender ? 

## 3.1.5 Are there difference in student performance between gender ? 
