---
title: "Hands-on_Ex5b"
date: "February 5,2024"
date-modified: "last-modified"
execute: 
  eval: true 
  echo: true
  warning: false
editor: visual 
---

# 6. Visual Correlation Analysis

## 6.1 Overview

Correlation coefficient is a popular statistic that use to measure the type and strength of the relationship between two variables. 

The values of a correlation coefficient ranges between -1.0 and 1.0. A correlation coefficient of 1 shows a perfect linear relationship between the two variables, while a -1.0 shows a perfect inverse relationship between the two variables. 

Reasons for computing a correlation matrix

-   To reveal the relationship between high-dimensional variables pair-wisely.

<!-- -->

-   To input into other analyses

-   As a diagnostic when checking other analyses

[Corrgram](http://www.datavis.ca/papers/corrgram.pdf) tend to be used to visually explore and analyse the structure and the patterns of relations among variables

Two main schemes:

-   Rendering the value of a correlation to depict its sign and magnitude, and

-   Reordering the variables in a correlation matrix so that "similar" variables are positioned adjacently, facilitating perception.

## 6.2 Loading R packages

```{r}
pacman::p_load(corrplot, ggstatsplot, tidyverse)
```

## 6.3 Preparing Data

### 6.3.1 Importing data

```{r}
wine <- read_csv("data/wine_quality.csv")
```

## 6.4 Building correlation matrix: pairs() method

## 6.4.1 Building a basic correlation matrix

```{r}
pairs(wine[,1:11])
```

It uses the default *pairs* function. Columns 2 to 12 of wine dataframe is used to build the scatterplot matrix. The variables are: fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates and alcohol.

```{r}
pairs(wine[,2:12])
```

### 6.4.2 Drawing the lower corner

To show the lower half of the correlation matrix, the upper.panel argument will be used as shown in the code chunk below.

```{r}
pairs(wine[,2:12], upper.panel = NULL)
```

Similarly, you can display the upper half of the correlation matrix by using the code chun below.

```{r}
pairs(wine[,2:12], lower.panel = NULL)
```

### 6.4.3 Including with correlation coefficients

To show the correlation coefficient of each pair of variables instead of a scatter plot, [*panel.cor*](https://www.rdocumentation.org/packages/xcms/versions/1.48.0/topics/panel.cor) function will be used. This will also show higher correlations in a larger font.

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,2:12], 
      upper.panel = panel.cor)
```

## 6.5 Visualizing correlation matrix: ggcormat()

One of the major limitation of the correlation matrix is that the scatter plots appear very cluttered when the number of observations is relatively large (i.e. more than 500 observations). To over come this problem, **Corrgram** data visualisation technique suggested

### 6.5.1 The basic plot

Advantage of using ggcorrmat() over many other method to visualize a correlation matrix is its ability to provide a comprehensive and yet professional statistical report

```{r}
ggstatsplot::ggcorrmat( data = wine,  cor.vars = 1:11)
```

```{r}
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  title    = "Correlogram for wine dataset",
  subtitle = "Four pairs are no significant at p < 0.05"
)
```

-   `cor.vars` argument is used to compute the correlation matrix needed to build the corrgram.

-   `ggcorrplot.args` argument provide additional (mostly aesthetic) arguments that will be passed to [`ggcorrplot::ggcorrplot`](http://www.sthda.com/english/wiki/ggcorrplot-visualization-of-a-correlation-matrix-using-ggplot2) function

```{r}
ggplot.component = list(
    theme(text=element_text(size=5),
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8)))

```

## 6.6 Building multiple plots

```{r}
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```

-   to build a facet plot, the only argument needed is `grouping.var`.

-   Behind *group_ggcorrmat()*, **patchwork** package is used to create the multiplot. `plotgrid.args` argument provides a list of additional arguments passed to [*patchwork::wrap_plots*](https://patchwork.data-imaginist.com/reference/wrap_plots.html), except for guides argument which is already separately specified earlier.

-   Likewise, `annotation.args` argument is calling [*plot annotation arguments*](https://patchwork.data-imaginist.com/reference/plot_annotation.html) of patchwork package.

## 6.7 Visualizing correlation matrix using corrplot package

### 6.7.1 Getting started with corrplot

 [*cor()*](https://www.rdocumentation.org/packages/stats/versions/3.6.0/topics/cor) of R Stats is used to compute the correlation matrix of wine data frame.

```{r}
wine.cor <- cor(wine[, 1:11])
```

Next, [*corrplot()*](https://www.rdocumentation.org/packages/corrplot/versions/0.2-0/topics/corrplot) is used to plot the corrgram by using all the default setting as shown in the code chunk below.

```{r}
corrplot(wine.cor)
```

### 6.7.2 Working with visual geometrics

In **corrplot** package, there are seven visual geometrics (parameter method) can be used to encode the attribute values. They are: circle, square, ellipse, number, shade, color and pie. The default is circle. As shown in the previous section, the default visual geometric of corrplot matrix is circle. However, this default setting can be changed by using the *method* argument as shown in the code chunk below.

```{r}
corrplot(wine.cor, 
         method = "ellipse") 
```

### 6.7.3 Working with layout

*corrplor()* supports three layout types, namely: "full", "upper" or "lower". The default is "full" which display full matrix.

The default setting can be changed by using the *type* argument of *corrplot()*.

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower")
```

After further customized:

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         type="lower",
         diag = FALSE,
         tl.col = "black")
```

### 6.7.4 Working with mixed layout

In order to create a coorgram with mixed layout, the [*corrplot.mixed()*](https://www.rdocumentation.org/packages/corrplot/versions/0.84/topics/corrplot.mixed), a wrapped function for mixed visualisation style will be used.

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

Notice that argument *lower* and *upper* are used to define the visualisation method used.

### 6.7.5 Combining corrgram with the significant test

In statistical analysis, we are also interested to know which pair of variables their correlation coefficients are statistically significant.

With corrplot package, we can use the *cor.mtest()* to compute the p-values and confidence interval for each pair of variables.

```{r}
wine.sig = cor.mtest(wine.cor, conf.level= .95)
```

```{r}
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)
```

### 6.7.6 Recorder a corrgram

Matrix reorder is very important for mining the hiden structure and pattern in a corrgram. 

Currently, **corrplot** package support four sorting methods, they are:

-   "AOE" is for the angular order of the eigenvectors. See Michael Friendly (2002) for details.

-   "FPC" for the first principal component order.

-   "hclust" for hierarchical clustering order, and "hclust.method" for the agglomeration method to be used.

    -   "hclust.method" should be one of "ward", "single", "complete", "average", "mcquitty", "median" or "centroid".

-   "alphabet" for alphabetical order.

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               order="AOE",
               tl.col = "black")
```

### 6.7.7 Reordering a correlation matrix using hclust

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         tl.pos = "lt",
         tl.col = "black",
         order="hclust",
         hclust.method = "ward.D",
         addrect = 3)
```
