---
title: "Univariate & Bivariate Analysis"
---

# Univariate & Bivariate Analysis

# 1. Overview

In this analysis, it will be break into two parts:

-   General visualization of student profile who fill in the questionnaire

-   EDA & CDA of related possible influential factors on students academic performance

# 2. Loading R package

```{r}
pacman::p_load(tidyverse, haven, patchwork,ggrepel,ggthemes, ggstatsplot, ggplot2, ggridges, performance, ggiraph)
```

# 3. Import data

```{r}
dataset_filtered <- read.csv("data/stu_SG_filtered.csv")
```

```{r}
dataset <- read.csv("data/stu_SG_rcd.csv")
```

# 4. Univariate Analysis

## 4.1 Student Profile: *Know more about our participants*

pie chart for gender, school type, home language

```{r}
dataset_gender <- dataset %>%
  count(Gender)

dataset_gender <- dataset_gender %>% 
  mutate(prop = n / sum(dataset_gender$n) *100,
         n_pct = paste0(round(prop,2),"%"),
         ypos = cumsum(prop)- 0.5*prop,
         ypos_pct = case_when(ypos > 50 ~ ypos-4, ypos<= 50 ~ ypos+4)) 
```

```{r}
ggplot(dataset_gender, aes(x="", y=prop, fill=Gender)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  theme_void() + 
  geom_text(aes(y = ypos, label = n), color = "white", size=3) +
  geom_text(aes(y = ypos_pct, label = n_pct), color = "black", size=3)


```

bar chart for immigration, home language, siblings, parents education

```{r}
ggplot(data = dataset,
       aes(x = Immigration)) +
  geom_bar() +
geom_text(stat="count", 
      aes(label=paste0(stat(count), ", ", 
      round(after_stat(count)/sum(after_stat(count))*100, 1), "%")),
      vjust=-0.5,
      size=3) +
    labs(x = "",
         y = "",
         title = "Immigration") + 
    theme(
      plot.title = element_text(size=9, face="bold"),
      panel.grid = element_blank(), 
      axis.text.y = element_blank(),
      axis.title.y = element_blank(),
      axis.ticks.y = element_blank(),   
      axis.ticks.x = element_blank(),      
      plot.background=element_rect(fill="white",colour="white")
    ) 
```

## 4.2 Academic performance

```{r}
# min, max, mean, median of the score 
math_mean <- round(mean(dataset$Math),0)
math_median <- round(median(dataset$Math),0)
math_min <- round(min(dataset$Math),0)
math_max <- round(max(dataset$Math),0)
leftwhisk_math <- round(quantile(dataset$Math, probs = .25)-1.5*IQR(dataset$Math),0)

math_distribution <- 
  ggplot(
    dataset, 
    aes(x = Math)
  ) +
  geom_histogram(
    bins = 25,
    color = "#f5f5f5",
    fill="#c7c8cc"
  ) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 5), 
    limits = c(150,950),) +
  geom_vline(
    aes(xintercept = math_median),
    colour="#595DE5", 
    linewidth = 0.7, 
    linetype = "dashed"
  ) +
  annotate(
    geom = "text", 
    x = 660, 
    y = 870,
    label = paste0("Median = ", math_median),
    color="#595DE5"
  ) +
  geom_vline(
    aes(xintercept = math_mean),
    colour="#e0218a", 
    linewidth = 0.7, 
    linetype = "dashed"
  ) +
  annotate(
    geom = "text", 
    x = 535, 
    y = 870,
    label = paste0("Mean = ", math_mean),
    color="#e0218a"
  ) +  
  labs(x = NULL, y = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill="#f5f5f5",colour="#f5f5f5")
  )

# Boxplot
box_math <- 
  ggplot(
    dataset, 
    aes(y = Math, x="")
  ) + 
  geom_boxplot(
    fill="#c7c8cc",
    outlier.colour="#d18a7d", 
    outlier.shape=16,
    outlier.size=1, 
    notch=FALSE
  ) + 
  stat_summary(geom = 'point',
               fun.y='mean',
               colour = '#e0218a',
               size=4) +
  coord_flip() + 
  
    #Annotate with amounts to show range
  annotate(
    geom = "text", 
    x = 0.5, 
    y = math_min,
    label = paste0(math_min),
    color = "#d18a7d"
  ) + 
   annotate(
    geom = "text", 
    x = 0.5, 
    y = leftwhisk_math,
    label = paste0(leftwhisk_math),
    color = "#34414E"
  ) +
  annotate(
    geom = "text", 
    x = 0.5, 
    y = math_max,
    label = paste0(math_max),
    color = "#34414E"
  ) + 
  
  labs(
    title = "Distribution of Math Scores Performance",
     y = "", 
    x = ""
  ) + 
   scale_y_continuous(
    limits = c(150,950), 
  ) +
  
  stat_boxplot(
    geom="errorbar", 
    width=0.5
  ) + 
  theme_minimal() +
  theme(
    axis.text = element_blank(), 
    axis.ticks = element_blank(),
    panel.grid.major = element_blank(),
     plot.background = element_rect(
       fill="#f5f5f5",colour="#f5f5f5")
  )

hbox_math <- (box_math / math_distribution) + 
        plot_layout(heights = c(1, 2.5))

hbox_math & theme_void()

  

```

# 5 Bivariate Analysis

## 5.1 Time spend on homework & Academic performance

### 5.1.1 Distribution

```{r}
ggplot(
    dataset, 
    aes(x = Math, 
        y = Homework_Math,
        fill = Homework_Math,
        color = Homework_Math)
  ) +
  geom_density_ridges(
    # Adjust bandwidth to reveal variations 
    bandwidth = 15,
    alpha = .6
  ) +
  labs(
    title = "Distribution of Math Score Across Time Length Spend on Math Homework"
  ) +
  theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill="white",colour="white")
  )
```

### 5.1.2 Normality Test

```{r}
model1 <- lm(Math ~ Homework_Math, data = dataset)
check_n <- check_normality(model1)
plot(check_n)
```

### 5.1.3 ANOVA

```{r}
ggbetweenstats(
  data = dataset,
  x = Homework_Math, 
  y = Math,
  type = "non-parametric",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

## 5.2 Exercise time & Academic performance

### 5.2.1 Distribution

```{r}
ggplot(dataset, 
    aes(x = Math, 
        y = Preference_Math)
  ) +
  geom_point(
    alpha = 0.2, 
    size = 7
  ) +
  stat_summary(
    color = "#fdbf11", 
    fun = "median", 
    geom = "point", 
    size = 7, 
    alpha = 0.8
  ) +
  scale_x_continuous(
    expand = expansion(mult = c(0.02, 0)), 
    limits = c(150, 950),
    breaks = scales::pretty_breaks(n=3)
  ) +
  scale_y_discrete(
    limits = c(NA,"Strongly Agree","Agree","Disagree","Strongly Disagree")
  ) +
  labs(
    title = "Stronger perference in math tend to have a higher math score ",
    subtitle = "Median values are highlighted in Yellow",
    x = "Math Score",
    y = NULL
  ) +
  theme(
    axis.ticks.y = element_blank(),
    plot.background = element_rect(fill="white",colour="white")
  )
```

### 5.2.2 Normality Test

```{r}
model1 <- lm(Math ~ Preference_Math, data = dataset)
check_n <- check_normality(model1)
plot(check_n)
```

### 5.2.3 ANOVA

```{r}
ggbetweenstats(
  data = dataset,
  x = Preference_Math, 
  y = Math,
  type = "non-parametric",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

## 5.3 Academic performance across number of books and gender 

```{r}
# Calculate mean joviality per tile
gbcols <- c("Gender", "Books")

# Calculate Mean joviality score by agegroup and incomegroup and save in new dataframe
Math_performance <- dataset %>% 
  group_by(
    across(all_of(gbcols))) %>% 
  # Create new variable for mean joviality score by both groups
  summarise(
    mean_math = round(mean(Math),2), 
    .groups = "drop") %>%
  ungroup()

# Style tooltip
tooltip_css <- "background-color:#3E4254; font-style:bold; color:#F8F3E6;" 

# Tile Plot
gg_tile <-
  ggplot(
    Math_performance,
    aes(x = Gender, 
        y = Books)
  ) +
  geom_tile_interactive(
    aes(fill = mean_math,
        # Format tooltip information
        tooltip = paste("Gender: ", Gender, "<br>",
                        "Books: ", Books, "<br>",
                        "Mean Math Score: ", mean_math))
  ) + 
  # Specify gradient colors for divergent scale 
  scale_fill_gradient(
    low = "#0a4c6a",
    high = "#fed1c7",
    space = "Lab",
    na.value = "grey50",
    guide = "colourbar",
    aesthetics = "fill"
  ) +  
 labs(
   title = "Math Score across Gender and Number of Books",
    subtitle = "Younger People from lower Income Groups had the highest joviality scores",
    x = "Gender",
    y = "Books"
  ) +
  theme(
    # Move legend to right and change orientation so values are not overlapping 
    legend.position = "right",
    legend.direction = "vertical",
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    text = element_text(family = "Lato"),
    plot.background = element_rect(
      fill="white",colour="white")
  ) + 
  scale_y_discrete(
    limits = c(NA,"0","1 - 10","11 - 25","26 - 100", "101 - 200", "201-500", ">500"))

# Make into interactive ggiraph object
ggiraph::girafe(ggobj = gg_tile,
                width_svg = 6,
                height_svg = 6*0.7,
                options = list(opts_tooltip(css = tooltip_css)))                   
```

## 5.4 Time of Exercise and Academic performance 
