---
title: "Hands-on_Ex7"
date: "February 6,2024"
date-modified: "last-modified"
execute: 
  eval: true 
  echo: true
  warning: false
editor: visual 
---

# Choropleth Mapping with R

# 21.1 Overview 

Choropleth mapping involves the symbolisation of enumeration units, such as countries, provinces, states, counties or census units, using area patterns or graduated colors.

# 21.2 Getting Started 

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

# 21.3 Importing Data into R

## 21.3.1 The Data 

## 21.3.2 Importing Geospatial Data into R

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")

```

Examine the content of mpsz by using code chunk below

```{r}
mpsz
```

## 21.3.3 Importing attribute data into R

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

## 21.3.4 Data preparation 

The data table should include the variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.

-   YOUNG: age group 0 to 4 until age groyup 20 to 24,

-   ECONOMY ACTIVE: age group 25-29 until age group 60-64,

-   AGED: age group 65 and above,

-   TOTAL: all age group, and

-   DEPENDENCY: the ratio between young and aged against economy active group

### 21.3.4.1 Data wrangling

-   *pivot_wider()* of **tidyr** package, and

-   *mutate()*, *filter()*, *group_by()* and *select()* of **dplyr** package

```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

### 21.3.4.2 Joining the attribute data and geospatial data

```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

*left_join()* of **dplyr** package is used with `mpsz` simple feature data frame as the left data table is to ensure that the output will be a simple features data frame.

```{r}
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

# 21.4 Choropleth mapping geospatial data using tmap

Two approaches can be used to prepare thematic map using *tmap*, they are:

-   Plotting a thematic map quickly by using *qtm()*.

-   Plotting highly customisable thematic map by using tmap elements.

```{r}
tmap_mode("plot") # tmap_mode() with plot option is used to produce a static map
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY") #fill argumemt is used to map the attribute 
```

## 21.4.2 Creating a choropleth map by using tmap's elements

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

In the following sub-section, we will share with you tmap functions that used to plot these elements.

### 21.4.2.1 Drawing a base map

The basic building block of **tmap** is 

-   *tm_shape()* followed by one or more layer elemments such as *tm_fill()* and *tm_polygons()*.

```{r}
tm_shape(mpsz_pop2020) +  #define the input data 
  tm_polygons() # draw the planning subzone polygons 
```

### 21.4.2.2 Drawing a choropleth map using tm_polygons ()

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY") # target variable 
```

### 21.4.2.3 Drawing a choropleth map using tm_fill() and tm_border()

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

To add the boundary of the planning subzones, tm_borders will be used

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1) # alpha argument is used to define transparency number between 0 (totally transparent) and 1 (not transparent)
```

Beside *alpha* argument, there are three other arguments for *tm_borders()*, they are:

-   *col* = border colour,

-   *lwd* = border line width. The default is 1, and

-   *lty* = border line type. The default is \"solid\".

## 21.4.3 Data classification methods of tmap

**tmap** provides a total ten data classification methods, namely: *fixed*, *sd*, *equal*, *pretty* (default), *quantile*, *kmeans*, *hclust*, *bclust*, *fisher*, and *jenks*.

To define a data classification method, the *style* argument of *tm_fill()* or *tm_polygons()* will be used.

### 21.4.3.1 Plotting choropleth maps with built-in classification methods 

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

Equal data classification method

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

Distribution of quantile data classification method are more evenly distributed then equal data classification method

Different classification methods

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "fisher") +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "hclust") +
  tm_borders(alpha = 0.5)
```

Similar classification method and different numbers of classes

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 2,
          style = "hclust") +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "hclust") +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 10,
          style = "hclust") +
  tm_borders(alpha = 0.5)
```

### 21.4.3.2 Plotting choropleth map with custome break

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

## 21.4.4 Color scheme 

### 21.4.4.1 Using colorbrewer palette 

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

To reverse the color shading, add a "-" prefix

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```

## 21.4.5 Map Layouts 

Map layout refers to the combination of all map elements into a cohensive map. Map elements include among others the objects to be mapped, the title, the scale bar, the compass, margins and aspects ratios. Colour settings and data classification methods covered in the previous section relate to the palette and break-points are used to affect how the map looks.

### 21.4.5.1 Map Legend 

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

### 21.4.5.2 Map Style 

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

### 21.4.5.3 Cartographic Furniture 

*tm_compass()*, *tm_scale_bar()* and *tm_grid()* are used to add compass, scale bar and grid lines onto the choropleth map.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

```{r}
tmap_style("white")
```

## 21.4.6 Drawing Small Multiple Choropleth maps 

**Small multiple maps**, also referred to as **facet maps**, are composed of many maps arrange side-by-side, and sometimes stacked vertically. Small multiple maps enable the visualisation of how spatial relationships change with respect to another variable

In **tmap**, small multiple maps can be plotted in three ways:

-   by assigning multiple values to at least one of the asthetic arguments,

-   by defining a group-by variable in *tm_facets()*, and

-   by creating multiple stand-alone maps with *tmap_arrange()*.

### 21.4.6.1 By assigning multiple values to at least one of the aesthetic arguments

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

```{r}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

### 21.4.62 By defining a group-by variable in tm_facets ()

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=FALSE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

### 21.4.6.3 By creating multiple stand-alone maps with tmap_arrange()

```{r}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

## 21.4.7 Mapping spatial object meeting a selection criterion

```{r}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

The current graph its only one certain region in singapore
