---
title: "Take-home_Ex3"
date: "February 11,2024"
date-modified: "last-modified"
execute: 
  eval: true 
  echo: true
  warning: false
editor: visual 
---

# Overview: Climate Change in Singapore

-   An increase of daily mean temperature

-   The temperature difference between wet and dry month is predicted to increase

# Objective

Using interactive techniques to demonstrate data visualization of monthly temperature in Singapore between year 1983, 1993, 2003, 2013, 2023.

# Data Preparation

## 1.1 Loading R package

Let's load the libraries and R package needed for this assignment using `pacman::p_load()` function:

```{r}
pacman::p_load(tidyverse, )
```

The loaded packages are:

-   [tidverse](https://www.tidyverse.org/): data wrangling, tidying, reading/writing, parsing, and visualizing

## 1.2 Dataset

The dataset using in this assignment is taking from historical daily temperature or rainfall data in Meteorological service Singapore website.

### 1.2.1 Import the datasets

In this assignment, we will using data of daily temperature in Changi region from July of 1983, 1993, 2003, 2013, to 2023.

```{r}
period1<-read_csv("data/DAILYDATA_S24_202307.csv")
```

```{r}
period2<-read_csv("data/DAILYDATA_S24_201307.csv")
```

```{r}
perid3<-read_csv("data/DAILYDATA_S24_200307.csv")
```

```{r}
perid4<-read_csv("data/DAILYDATA_S24_199307.csv")
```

```{r}
perid5<-read_csv("data/DAILYDATA_S24_198307.csv")
```

# Visualization Techniques Selection

# Data Visualization
