---
title: "Hands-on_Ex8"
date: "March 4,2024"
date-modified: "last-modified"
execute: 
  eval: true 
  echo: true
  warning: false
  freeze: true
editor: visual 
---

# Modeling, Visualizing and Analyzing Network Data with R

# 27.1 Overview

Main target: model, analyze and visualize network data

-   create graph object data frames

-   build network graph visualization

-   compute network geometrics

-   build advanced graph visualization by incorporating the network geometrics

-   build interactive network visualization

# 27.2 Getting Started 

## 27.2.1 Loading R packages 

```{r}
pacman::p_load(igraph, tidygraph, ggraph, 
               visNetwork, lubridate, clock,
               tidyverse, graphlayouts)
```

# 27.3 The data

## 27.3.3 Import network data 

```{r}
GAStech_nodes <- read_csv("data/GAStech_email_node.csv")
GAStech_edges <- read_csv("data/GAStech_email_edge-v2.csv")
```

## 27.3.4 Reviewing the imported data 

```{r}
glimpse(GAStech_edges)
```

### 27.3.5 Wrangling time 

```{r}
GAStech_edges <- GAStech_edges %>%
  mutate(SendDate = dmy(SentDate)) %>%
  mutate(Weekday = wday(SentDate,
                        label = TRUE,
                        abbr = FALSE))
```

Things to learn from the code chunk above

-   *dmy()* transforms the SentDate to Date data type.

-   *wday()* returns the day of the week as a decimal number or an ordered factor if label is TRUE. The argument abbr is FALSE keep the daya spells in full, i.e. Monday. The function will create a new column in the data.frame i.e. Weekday and the output of *wday()* will save in this newly created field.

-   the values in the *Weekday* field are in ordinal scale.

## 27.3.7 Wrangling attributes 

Aggregate the individual by date, senders, receivers, main subject and day of the week

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject == "Work related") %>%
  group_by(source, target, Weekday) %>%
    summarise(Weight = n()) %>%
  filter(source!=target) %>%
  filter(Weight > 1) %>%
  ungroup()
```

## 27.3.7 Reviewing the revised edges file 

```{r}
glimpse(GAStech_edges_aggregated)
```

# 27.4 Creating network objects using tidygraph 

It provides a tidy API for graph/network manipulation. While network data itself is not tidy, it can be envisioned as two tidy tables, one for node data and one for edge data. tidygraph provides a way to switch between the two tables and provides dplyr verbs for manipulating them

## 27.4.1 The tbl_graph object 

-   [`tbl_graph()`](https://tidygraph.data-imaginist.com/reference/tbl_graph.html) creates a **tbl_graph** network object from nodes and edges data.

-   [`as_tbl_graph()`](https://tidygraph.data-imaginist.com/reference/tbl_graph.html) converts network data and objects to a **tbl_graph** network.

    Below are network data and objects supported by `as_tbl_graph()`

    -   a node data.frame and an edge data.frame,

    -   data.frame, list, matrix from base,

    -   igraph from igraph,

    -   network from network,

    -   dendrogram and hclust from stats,

    -   Node from data.tree,

    -   phylo and evonet from ape, and

    -   graphNEL, graphAM, graphBAM from graph (in Bioconductor).

## 27.4.1 The dplyr verbs in tidygraph 

-   *activate()* verb from **tidygraph** serves as a switch between tibbles for nodes and edges.

-   In the above the *.N()* function is used to gain access to the node data while manipulating the edge data. Similarly *.E()* will give you the edge data and *.G()* will give you the **tbl_graph** object itself.

## 27.4.3 Using `tbl_graph()` to build tidygraph data model

```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated, 
                           directed = TRUE)
```

## 27.4.4 Reviewing the output tidygraph\'s graph object

```{r}
GAStech_graph
```

## 27.4.5 Reviewing the output tidygraph\'s graph object

-   The output above reveals that *GAStech_graph* is a tbl_graph object with 54 nodes and 4541 edges.

-   The command also prints the first six rows of \"Node Data\" and the first three of \"Edge Data\".

-   It states that the Node Data is **active**. The notion of an active tibble within a tbl_graph object makes it possible to manipulate the data in one tibble at a time.

## 27.4.6 Changing the active object

The nodes tibble data frame is activated by default, but you can change which tibble data frame is active with the *activate()* function. Thus, if we wanted to rearrange the rows in the edges tibble to list those with the highest \"weight\" first, we could use *activate()* and then *arrange()*.

```{r}
GAStech_graph %>%
  activate(edges) %>%
  arrange(desc(Weight))
```

# 27.5 Plotting static network graphs with ggraph package 

As in all network graph, there are three main aspects to a **ggraph**\'s network graph, they are:

-   [nodes](https://cran.r-project.org/web/packages/ggraph/vignettes/Nodes.html),

-   [edges](https://cran.r-project.org/web/packages/ggraph/vignettes/Edges.html) and

-   [layouts](https://cran.r-project.org/web/packages/ggraph/vignettes/Layouts.html).

## 27.5.1 Plotting a basic network graph 

```{r}
ggraph(GAStech_graph) +
  geom_edge_link() +
  geom_node_point()
```

Things to learn from the code chunk above

-   The basic plotting function is `ggraph()`, which takes the data to be used for the graph and the type of layout desired. Both of the arguments for `ggraph()` are built around *igraph*. Therefore, `ggraph()` can use either an *igraph* object or a *tbl_graph* object.

## 27.5.2 Changing the default network graph theme 

```{r}
g <- ggraph(GAStech_graph) + 
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

Things to learn from the code chunk above

-   **ggraph** introduces a special ggplot theme that provides better defaults for network graphs than the normal ggplot defaults. `theme_graph()`, besides removing axes, grids, and border, changes the font to Arial Narrow (this can be overridden).

-   The ggraph theme can be set for a series of plots with the `set_graph_style()` command run before the graphs are plotted or by using `theme_graph()` in the individual plots.

## 27.5.3 Changing the coloring of the plot

```{r}
g <- ggraph(GAStech_graph) + 
  geom_edge_link(aes(colour = 'grey50')) +
  geom_node_point(aes(colour = 'grey40'))

g + theme_graph(background = 'grey10',
                text_colour = 'white')
```

## 27.5.4 Working with ggraph's layouts

**ggraph** support many layout for standard used, they are: star, circle, nicely (default), dh, gem, graphopt, grid, mds, spahere, randomly, fr, kk, drl and lgl. Figures below and on the right show layouts supported by `ggraph()`.

## 27.5.5 Fruchterman and reingold layout 

```{r}
g <- ggraph(GAStech_graph, 
            layout = "fr") +
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

## 27.5.6 Modifying network nodes

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes()) +
  geom_node_point(aes(colour = Department, 
                      size = 3))

g + theme_graph()
```

Things to learn from the code chunks above:

-   *geom_node_point* is equivalent in functionality to *geo_point* of **ggplot2**. It allows for simple plotting of nodes in different shapes, colours and sizes. In the codes chnuks above colour and size are used.

## 27.5.7 Modifying edges 

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") +
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 3)

g + theme_graph()
```

Things to learn from the code chunks above:

-   *geom_edge_link* draws edges in the simplest way - as straight lines between the start and end nodes. But, it can do more that that. In the example above, argument *width* is used to map the width of the line in proportional to the Weight attribute and argument alpha is used to introduce opacity on the line.

## 27.6 Creating facet graphs 

There are three functions in ggraph to implement faceting, they are:

-   [*facet_nodes()*](https://r4va.netlify.app/chap27) whereby edges are only draw in a panel if both terminal nodes are present here,

-   [*facet_edges()*](https://ggraph.data-imaginist.com/reference/facet_edges.html) whereby nodes are always drawn in al panels even if the node data contains an attribute named the same as the one used for the edge facetting, and

-   [*facet_graph()*](https://ggraph.data-imaginist.com/reference/facet_graph.html) faceting on two variables simultaneously.

## 27.6.1 Working with facet_edges()

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)

g + facet_edges(~Weekday)
```

## 27.6.2 Working with facet_edges()

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2) +
  theme(legend.position = 'bottom')
  
g + facet_edges(~Weekday)
```

## 27.6.3 A framed facet graph

```{r}
set_graph_style() 

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_edges(~Weekday) +
  th_foreground(foreground = "grey80",  
                border = TRUE) +
  theme(legend.position = 'bottom')
```

## 27.6.4 Working with facet_nodes()

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_nodes(~Department)+
  th_foreground(foreground = "grey80",  
                border = TRUE) +
  theme(legend.position = 'bottom')
```

# 27.7 Network Metrics analysis 

## 27.7.1 Computing centrality indices

Centrality measures are a collection of statistical indices use to describe the relative important of the actors are to a network.

There are four well-known centrality measures, namely: degree, betweenness, closeness and eigenvector. 

```{r}
g <- GAStech_graph %>%
  mutate(betweenness_centrality = centrality_betweenness()) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department,
            size=betweenness_centrality))
g + theme_graph()
```

Things to learn from the code chunk above:

-   *mutate()* of **dplyr** is used to perform the computation.

-   the algorithm used, on the other hand, is the *centrality_betweenness()* of **tidygraph**.

## 27.7.2 Visualizing network metrics 

```{r}
g <- GAStech_graph %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department, 
                      size = centrality_betweenness()))
g + theme_graph()
```

## 27.7.3 Visualizing community 

tidygraph package inherits many of the community detection algorithms imbedded into igraph and makes them available to us, including *Edge-betweenness (group_edge_betweenness)*, *Leading eigenvector (group_leading_eigen)*, *Fast-greedy (group_fast_greedy)*, *Louvain (group_louvain)*, *Walktrap (group_walktrap)*, *Label propagation (group_label_prop)*, *InfoMAP (group_infomap)*, *Spinglass (group_spinglass)*, and *Optimal (group_optimal)*. Some community algorithms are designed to take into account direction or weight, while others ignore it. Use this [link](https://tidygraph.data-imaginist.com/reference/group_graph.html) to find out more about community detection functions provided by tidygraph,

```{r}
g <- GAStech_graph %>%
  mutate(community = as.factor(group_edge_betweenness(weights = Weight, directed = TRUE))) %>%
  ggraph(layout = "fr") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = community))  

g + theme_graph()
```

# 27.8 Building interactive network graph with visNetwork

-   *visNetwork()* function uses a nodes list and edges list to create an interactive graph.

    -   The nodes list must include an \"id\" column, and the edge list must have \"from\" and \"to\" columns.

    -   The function also plots the labels for the nodes, using the names of the actors from the \"label\" column in the node list.

-   The resulting graph is fun to play around with.

    -   You can move the nodes and the graph will use an algorithm to keep the nodes properly spaced.

    -   You can also zoom in and out on the plot and move it around to re-center it.

## 27.8.1 Data Preparation

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  left_join(GAStech_nodes, by = c("sourceLabel" = "label")) %>%
  rename(from = id) %>%
  left_join(GAStech_nodes, by = c("targetLabel" = "label")) %>%
  rename(to = id) %>%
  filter(MainSubject == "Work related") %>%
  group_by(from, to) %>%
    summarise(weight = n()) %>%
  filter(from!=to) %>%
  filter(weight > 1) %>%
  ungroup()
```

## 27.8.2 Plotting the first interactive network graph

```{r}
visNetwork(GAStech_nodes, 
           GAStech_edges_aggregated)
```

## 27.8.3 Working with layout

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") 
```

## 27.8.4 Working with visual attributes- nodes

```{r}
GAStech_nodes <- GAStech_nodes %>%
  rename(group = Department) 
```

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

## 27.8.5 Working with visual attributes- Edges

In the code run below *visEdges()* is used to symbolise the edges.\
- The argument *arrows* is used to define where to place the arrow.\
- The *smooth* argument is used to plot the edges using a smooth curve.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visEdges(arrows = "to", 
           smooth = list(enabled = TRUE, 
                         type = "curvedCW")) %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```

## 27.8.6 Interactivity

In the code chunk below, *visOptions()* is used to incorporate interactivity features in the data visualisation.

-   The argument *highlightNearest* highlights nearest when clicking a node.

-   The argument *nodesIdSelection* adds an id node selection creating an HTML select element.

```{r}
visNetwork(GAStech_nodes,
           GAStech_edges_aggregated) %>%
  visIgraphLayout(layout = "layout_with_fr") %>%
  visOptions(highlightNearest = TRUE,
             nodesIdSelection = TRUE) %>%
  visLegend() %>%
  visLayout(randomSeed = 123)
```
